---
title: "class19"
author: "Hyeonseok Jang (PID# A59011126)"
date: "12/3/2021"
output:
  html_document: default
  pdf_document: default
---

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale.

Now, I want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378…) on ORMDL3 expression.


> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
```

```{r}
table(expr$geno)
```
The sample sizes for each genotypes are 108(A/A), 233(A/G), and 121(G/G).

```{r}
summary(expr[expr$geno=="A/A",]$exp)
summary(expr[expr$geno=="A/G",]$exp)
summary(expr[expr$geno=="G/G",]$exp)
```
The median expression levels for each of these genotypes are 31.25(A/A), 25.065(A/G), and 20.074(G/G).


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)

ggplot(expr) + aes(geno, exp, fill=geno) +
  geom_boxplot(outlier.shape=NA, notch=TRUE) +
  geom_jitter(width=0.2, alpha=0.2) +
  xlab("Genotype") + ylab("Expression") +
  theme(legend.position="none")
```
I can see clear difference in expression value between A/A and G/G. Based on the observation, I think that having a G/G is associated with the reduced expression of the gene.