---
title: "class10_Halloween_mini_project"
author: "Hyeonseok Jang (PID# A59011126)"
date: "10/29/2021"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

## 1. Importing candy data

```{r}
candy_file <- "candy-data.txt"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

## 2. What is your favorate candy?

```{r}
candy["Twix", ]$winpercent
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Skittles original", ]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

```{r}
#install.packages("skimr")
library("skimr")
skim(candy)
```
Winpercent is on a different scale to the majority of the other variables that are on a 0 to 1 scale.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

0 means that the candy has no chocolate in it, and  1 means that it does not.

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

No

> Q10. Is the center of the distribution above or below 50%?

It is below 50%.

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])>mean(candy$winpercent[as.logical(candy$fruity)])
```
Chocolate candy is on average higher ranked than fruit candy.

> Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruity)])
```
Yes, the difference is statistically significant.

## 3. Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```

Now, for the first time, using this plot we can answer questions like:

> Q17. What is the worst ranked chocolate candy?

It is Charlseton Chew.

> Q18. What is the best ranked fruity candy?

It is Starburst.

## 4. Taking a look at pricepercent

```{r}
#install.packages("ggrepel")
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

It is Reeses Miniatures.

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head(candy[ord,c(11,12)], n=5)
```
Nik L Nip is the least popular among the top 5 most expensive candy types in the dataset.

## 5. Exploring the correlation structure

```{r}
#install.packages("corrplot")
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity and chocolate

> Q23. Similarly, what two variables are most positively correlated?

Winpercent and chocolate

## 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, size=winpercent/100, text=rownames(my_data), label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space", subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)", caption="Data from 538")
```

```{r}
library(plotly)
```

```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus are picked up strongly by PC1 in the positive direction. It makes sense to me since they are positively correlated each other.

