---
title: "class08"
author: "Hyeonseok Jang (PID# A59011126)"
date: "10/22/2021"
output: github_document
---
#Clustering methods

kmeans clustering in R is done with the 'kmeans()' function.
Here we makeup some data to test and learn with.

```{r}
tmp <- c(rnorm(30,3), rnorm(30,-3))
data <- cbind(x=tmp, y=rev(tmp))
plot(data)
```

Run 'kmeans()' set k to 2 nstart 20. The thing with kmeans is you have to tell it how many clusters you want.

```{r}
km <- kmeans(data, centers = 2, nstart = 20)
km
```

> Q. How many point are in each cluster?

```{r}
km$size
```

> Q. What 'component' of your result object details cluster assignment/membership?

```{r}
km$cluster
```

> Q. What 'component' of your result object details cluster center?

```{r}
km$centers
```

> Q. Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(data, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```


# Hierarchical Clustering

We will use the 'hclust()' function on the same data as before and see how this method works.

```{r}
hc <- hclust(dist(data))
hc
```

hclust has a plot method

```{r}
plot(hc)
abline(h=7, col="red")
```

To find our membership vector we need to "cut" the tree and for this we use the 'cutree()' functino and tell it the height to cut at.

```{r}
cutree(hc, h=7)
```

We can also use 'cutree()' and take the number of x clusters we want...

```{r}
grps <- cutree(hc, h=7)
```

```{r}
plot(data, col=grps)
```


# Principal Component Analysis (PCA)

PCA is a super useful analysis method when you have lots of dimensions in your data...

##PCA of UK food data

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names=1)
x
```

How many rows and cols?
```{r}
dim(x)
```

```{r}
x
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
x
```

```{r}
barplot(as.matrix(x), col=rainbow(17), beside=TRUE)
```

```{r}
mycols <- rainbow(nrow(x))
pairs(x, col=mycols, pch=16)
```

##PCA to the rescue!

Here we will use the base R function for PCA, which is called 'prcomp()'. This function wants the transpose of data.

```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
plot(pca)
```

We want score plot (a.k.a. PCA plot).
```{r}
attributes(pca)
```

We are after the pca$x component for this plot...
```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```

We can also examine the PCA "loadings", which tell us how much the original variables contribute to each new PC...
```{r}
pca$rotation
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las = 2)
```


## One more PCA for today

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

```{r}
nrow(rna.data)
```

```{r}
ncol(rna.data)
```

```{r}
colnames(rna.data)
```

```{r}
pca.rna <- prcomp(t(rna.data), scale=TRUE)
summary(pca.rna)
```

```{r}
plot(pca.rna)
```

```{r}
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels = colnames(rna.data))
```