---
title: "class17"
author: "Hyeonseok Jang (PID# A59011126)"
date: "11/24/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

##Background
The goal of this hands-on mini-project is to examine and compare the Covid-19 vaccination rates around San Diego.

##Getting started
```{r}
#Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

> Q. How many entries do we have?

```{r}
nrow(vax)
```

> Q1. What column details the total number of people fully vaccinated?

```{r}
head(vax$persons_fully_vaccinated)
```

> Q2. What column details the Zip code tabulation area?

```{r}
head(vax$zip_code_tabulation_area)
```

> Q3. What is the earliest date in this dataset?

```{r}
head(sort(vax$as_of_date), 1)
```

> Q4. What is the latest date in this dataset?

```{r}
tail(sort(vax$as_of_date), 1)
```

We can use the **skimr** package and the 'skim()' function to get a quick overview of structure of this dataset.

```{r}
skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset?

9 columns

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
round(sum(is.na(vax$persons_fully_vaccinated))/nrow(vax)*100, 2)
```

##Working with dates

Notice that one of the columns is a data column. Working with the data we need **lubridate** package.
```{r}
##install.packages("lubridate")
library(lubridate)
```

This will not work because our data column was read as character..
```{r}
#today()-vax$as_of_date[1]
```

```{r}
d <- ymd(vax$as_of_date)
```

```{r}
today()-d[1]
```

I will make the 'as_of_date' column Date format..
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

> Q9. How many days have passed since the last update of the dataset?

```{r}
today()-vax$as_of_date[nrow(vax)]
```

> Q. How many days does the dataset span?

```{r}
vax$as_of_date[nrow(vax)]-vax$as_of_date[1]
```

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```

> Q. How many different ZIP code areas are in this dataset?

```{r}
length(unique(vax$zip_code_tabulation_area))
```

## Working with ZIP codes
To work with ZIP codes we can use the **zipcodeR**
```{r}
##install.packages("zipcodeR")
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

##Focus on the San Diego area

We want to subset the full CA 'vax' data down to just San Diego County.

We could do this with base R.
```{r}
sd <- vax[vax$county=="San Diego", ]
```

Subsetting can get tedious and complicated quickly when you have multiple things we want to subset by.
```{r}
##install.packages("dplyr")
library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

More complicated subsetting...
```{r}
sd.20 <- filter(vax, county=="San Diego",
      age5_plus_population > 20000)
nrow(sd.20)
```

> Q11. How many distinct zip codes are listed for San Diego County?

```{r}
length(unique(sd$zip_code_tabulation_area))
```

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-23”?

```{r}
sd.now <- filter(sd, as_of_date=="2021-11-23")
mean(sd.now$percent_of_population_fully_vaccinated, na.rm=TRUE)
summary(sd.now$percent_of_population_fully_vaccinated)
```

> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-23”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```
This plot above is going to be susceptible to being skewed by ZIP code areas with small population. This will have big effects for just a small number of unvax-ed folks..

##Focus on UCSD/La Jolla

```{r}
lj <- filter(sd, zip_code_tabulation_area=="92037")
```

> Q. What is the population of the 92037 ZIP code area?

```{r}
unique(lj$age5_plus_population)
```

> Q. What is the current vaccination value for this UCSD/La Jolla Zip code area?

```{r}
lj2 <- filter(sd.now, zip_code_tabulation_area=="92037")
lj2$percent_of_population_fully_vaccinated
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)
ljvaxrate <- ggplot(lj) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0, 1)) +
  labs(title="Vaccination rate for La Jolla CA 92037", x="Date", y="Percent Vaccinated")
ljvaxrate
```

##Comparing to similar sized areas
Let's make this plot for all San Diego County ZIP cod areas that have a population as least as large as 92037.
```{r}
sd.36 <- filter(sd, age5_plus_population > 36144)
head(sd.36)
```

> Q. How many ZIP code areas in San Diego county have a population larger than 92037?

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```

Lets make the plot.
```{r}
ggplot(sd.36) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line() +
  ylim(c(0, 1)) +
  labs(x="Date", y="Percent Vaccinated")
```

> Q. Make a plot like this for the all ZIP code areas in the State with a population at least as large as La Jolla.

```{r}
ca <- filter(vax, age5_plus_population > 36144)
length(unique(ca$zip_code_tabulation_area))
```

```{r}
ggplot(ca) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  ylim(c(0, 1)) +
  labs(x="Date", y="Percent Vaccinated")
```

> Q. What is the mean across the state for these 36k+ population areas?

```{r}
ca.now <- filter(ca, as_of_date=="2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ca) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  ylim(c(0, 1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept=mean(ca.now$percent_of_population_fully_vaccinated), linetype=2, color="red")
```